---
title: "Stone Complexity Index - R Statistics"
output:
  pdf_document: default
  html_notebook: default
  html_document:
    df_print: paged
---

This is the R notebook to update the statistics for the SCI project. 
## Housekeeping:
```{r}
# load common dataScience packages using pacman
if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, knitr, kableExtra, readstata13, mosaic, xtable, summarytools, qwraps2)
```

```{r global_options, include=FALSE}
# This code sets the global options for the notebook and its output
# replace the curly braces above with: {r global_options, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```




## Load the Data:
```{r}
stones <- read.dta13("~/Box Sync/Research-stone complexity/PercSA50.dta",
  generate.factors = TRUE, nonint.factors = TRUE)

```

### Convert to a tibble
```{r}
stones <- as.tibble(stones)
# str(stones)
```

## Data Preperation:
### Dealing with empty values, nulls, nas:
- Replace 999s with na
 - use na instead of null as it keeps a spot - not just empty - says there is no value here
 
```{r}
stones <- na_if(stones, 999)
```
 
### Factors

```{r}
# Factors
factors <- stones[, sapply(stones, is.factor)]
factorsNames <- names(factors)
# factorsNames

# Character that might need to be Factors
chars <- stones[, sapply(stones, is.character)]
charsNames <- names(chars)
#charsNames
```

Look at possible levels in character variables:
```{r eval=FALSE, include=FALSE}
stones %>% group_by(infra_or_supra) %>% summarise(count = n())
view(dfSummary(chars))
```

Convert Chr to Factors when appropriate using tidyverse 'forcats'

```{r}
gender_levels <- c("M", "F")
stones$gender <- factor(stones$gender, levels = gender_levels)
YesNo_levels <- c("Yes", "No")
stones$ammoniumacidurate <- factor(stones$ammoniumacidurate, levels = YesNo_levels)
stones$caox <- factor(stones$caox, levels = YesNo_levels)
stones$caphos <- factor(stones$caphos, levels = YesNo_levels)
stones$ct <- factor(stones$ct, levels = YesNo_levels)
stones$flex_nephroscope <- factor(stones$flex_nephroscope, levels = YesNo_levels)
stones$renal_pelvis <- factor(stones$renal_pelvis, levels = YesNo_levels)
stones$staghorn <- factor(stones$staghorn, levels = YesNo_levels)
stones$stent <- factor(stones$stent, levels = YesNo_levels)
stones$struvite <- factor(stones$struvite, levels = YesNo_levels)
stones$upper <- factor(stones$upper, levels = YesNo_levels)
stones$ureter <- factor(stones$ureter, levels = YesNo_levels)
access_site_levels <- c("radiology", "infra", "supra 12")
stones$infra_or_supra <- factor(stones$infra_or_supra, levels = access_site_levels)
```

### Create new variables:
any_prior_stone
any_prior_tx on the same kidney
tubeless if pcn_type == none
Note, need to use 
```{r}
# any_prior_stone
stones <- stones %>% mutate(any_prior_stone = if_else(number_previous_stones > 0, "Yes", "No")) 
stones$any_prior_stone <- factor(stones$any_prior_stone, levels = YesNo_levels)

# any_prior_tx
stones <- stones %>% mutate(any_prior_tx = if_else(prior_treatment == "none", "Yes", "No")) 
stones$any_prior_tx <- factor(stones$any_prior_tx, levels = YesNo_levels)
# tubeless
stones <- stones %>% mutate(tubeless = if_else(pcn_type == "None", "Yes", "No")) 
stones$tubeless <- factor(stones$tubeless, levels = YesNo_levels)
# any_special_anatomy
stones <- stones %>% mutate(any_special_anatomy = if_else(special_gu_anatomy == "none", "No", "Yes")) 
stones$any_special_anatomy <- factor(stones$any_special_anatomy, levels = YesNo_levels)

#just need to convert to factor
stones$uricacid <- factor(stones$uricacid, levels = YesNo_levels)
stones$cystine <- factor(stones$cystine, levels = YesNo_levels)
```
### Check these new variables
```{r}
stones_check <- select(stones, number_previous_stones, any_prior_stone, prior_treatment, any_prior_tx, tubeless, any_special_anatomy, residFrag, residual_frags, UAComp, uricacid, CystineComp, cystine, any_special_anatomy, SpecialAnatomy)
```

I later realized that stone length and stone width were charaters, because they 
had several entries that included multiple ints separated by a comma. The largest number was always first, so
I want to capture the chars prior to the firts comma:
```{r}
stones <- stones %>% mutate(stoneLength = sub(',.*', '', stone_length))
stones$stoneLength <- sub('o', '0', stones$stoneLength)
stones$stoneLength <- as.numeric(stones$stoneLength)
stones <- stones %>% mutate(stoneWidth = sub(',.*', '', stone_width))
stones$stoneWidth <- sub('o', '0', stones$stoneWidth)
stones$stoneWidth <- as.numeric(stones$stoneWidth)
regexTest <- select(stones, stone_length, stoneLength, stone_width, stoneWidth)
```

### Dates
Convert DOS to date:

```{r}
stones$dos <- as.Date(stones$dos, format = "%m/%d/%Y")
```

### reorder the variables to better match our purposes:
```{r}
PtData <- select(stones, mrn, age, gender,bmi, weight, any_prior_stone, co_morbids1, co_morbids2, co_morbids3, asa, any_special_anatomy)
StoneData <- select(stones, side, number_current_stones, stoneLength, stone_sa, stone_size, stone_volume, stoneWidth, radius, isoSphereSA, GCIndex, stone_location, upper, renal_pelvis, lower_pole, lowerPole, ureter, ammoniumacidurate, caox, caphos, cystine, struvite, uricacid, CystineComp, UAComp)
TxData <- select(stones,additional_stone_procedure, complications, ct, cxr, cxr_findings, ebl_ml, flex_nephroscope, infra_or_supra, length_of_stay, or_time, precr, postcr, stent, unitstransfused, residFrag, anotherProcedure, any_prior_tx)
stones <- select(stones, names(PtData), names(StoneData), names(TxData))
```

Refresh the dataframes by type now that we have cleaned and converted:
```{r}
# Factors
factors <- stones[, sapply(stones, is.factor)]
factorsNames <- names(factors)

# Characters
chars <- stones[, sapply(stones, is.character)]
charsNames <- names(chars)

# Integers or numbers
ints <- stones[, sapply(stones, is.numeric)]
intsNames <- names(ints)
# intsNames
```

## Statistics
### Summary Stats:

Realized there was a bmi of 175, checked the patient's file and
found it was really 25.7 so changed below:
```{r}
#replace the 175 BMI
stones[4,4] <- 25.7
```

```{r eval=FALSE, include=FALSE}
view(dfSummary(stones))
```


```{r eval=FALSE, include=FALSE}
descr(stones)
```
### Frequency tables

```{r eval=FALSE, include=FALSE}
sapply(factors, freq, report.nas = FALSE)
# freq(stones$gender, report.nas = FALSE)
# freq(stones$side, report.nas = FALSE)
# freq(stones$residFrag, report.nas = FALSE)


```

```{r}
GCIttest <- t.test(GCIndex ~ residFrag, data = stones)
GCIttestP <- round(GCIttest$p.value, digits = 3)
GCIttestP
```
```{r}
chisq.test(stones$gender, stones$residFrag, simulate.p.value = TRUE)
genTable <- table(stones$gender, stones$residFrag)
pval <- chisq.test(genTable)
pval$p.value
genTable
```



```{r}
ggplot(stones, aes(residFrag, GCIndex)) + geom_boxplot()
```
```{r include=FALSE}
list <- cat(names(stones), sep = "\n")
```

```{r}
median(stones$age)
IQR(stones$age)
median_iqr(stones$age)
```


```{r}
chisq.test(stones$any_special_anatomy, stones$residFrag, simulate.p.value = TRUE)
genTable <- table(stones$any_special_anatomy, stones$residFrag)
pval <- chisq.test(genTable)
pval$p.value
genTable
```

create table 1, with stats: START HERE****THIS Table has errors. for instance the Left Right part
I don't think using the stonesFree and stonesResid is correct?
```{r Table1, knitr.table.format="latex", results='asis'}
stonesFree <- stones %>% filter(residFrag == "No")
stonesResid <- stones %>% filter(residFrag == "Yes")

### AGE ###
medianAge <- median_iqr(stones$age, digits = getOption("qwraps2_frmt_digits", 1), show_n = "never", na_rm = TRUE)
medianAgeFree <- median_iqr(stonesFree$age, digits = getOption("qwraps2_frmt_digits", 1), show_n = "never", na_rm = TRUE)
medianAgeResid <- median_iqr(stonesResid$age, digits = getOption("qwraps2_frmt_digits", 1), show_n = "never", na_rm = TRUE)
statAge <- t.test(age ~ residFrag, data = stones)
statAgePval <- round(statAge$p.value, digits = 3)

### Men ###
men <- n_perc0(stones$gender == "M", na_rm = TRUE)
menFree <- n_perc0(stonesFree$gender == "M", na_rm = TRUE)
menResid <- n_perc0(stonesResid$gender == "M", na_rm = TRUE)
statMen <- chisq.test(stones$gender, stones$residFrag, simulate.p.value = TRUE)
statMenPval <- round(statMen$p.value, digits = 3)

### Weight ###
meanWeight <- mean_sd(stones$weight, digits = getOption("qwraps2_frmt_digits", 1), show_n = "never", denote_sd = "paren", na_rm = TRUE)
meanWeightFree <- mean_sd(stonesFree$weight, digits = getOption("qwraps2_frmt_digits", 1), show_n = "never", denote_sd = "paren", na_rm = TRUE)
meanWeightResid <- mean_sd(stonesResid$weight, digits = getOption("qwraps2_frmt_digits", 1), show_n = "never", denote_sd = "paren", na_rm = TRUE)
statWeight <- t.test(weight ~ residFrag, data = stones)
statWeightPval <- round(statWeight$p.value, digits = 3)

### BMI ###
meanBmi <- mean_sd(stones$bmi, digits = getOption("qwraps2_frmt_digits", 1), show_n = "never", denote_sd = "paren", na_rm = TRUE)
meanBmiFree <- mean_sd(stonesFree$bmi, digits = getOption("qwraps2_frmt_digits", 1), show_n = "never", denote_sd = "paren", na_rm = TRUE)
meanBmiResid <- mean_sd(stonesResid$bmi, digits = getOption("qwraps2_frmt_digits", 1), show_n = "never", denote_sd = "paren", na_rm = TRUE)
statBmi <- t.test(bmi ~ residFrag, data = stones)
statBmiPval <- round(statBmi$p.value, digits = 3)

### ASA ###
medianASA <-  median_iqr(stones$asa, digits = getOption("qwraps2_frmt_digits", 1), show_n = "never", na_rm = TRUE)
medianASAFree <-  median_iqr(stonesFree$asa, digits = getOption("qwraps2_frmt_digits", 1), show_n = "never", na_rm = TRUE)
medianASAResid <-  median_iqr(stonesResid$asa, digits = getOption("qwraps2_frmt_digits", 1), show_n = "never", na_rm = TRUE)
statASA <- t.test(asa ~ residFrag, data = stones)
statASAPval <- round(statASA$p.value, digits = 3)

### Prior Stones ###
priorStone <- n_perc0(stones$any_prior_stone == "Yes", na_rm = TRUE)
priorStoneFree <- n_perc0(stonesFree$any_prior_stone == "Yes", na_rm = TRUE)
priorStoneResid <- n_perc0(stonesResid$any_prior_stone == "Yes", na_rm = TRUE)
statPriorStone <- chisq.test(stones$any_prior_stone, stones$residFrag, simulate.p.value = TRUE)
statPriorStonePval <- round(statPriorStone$p.value, digits = 3)

### Special GU anatomy ###
specialAnat <- n_perc0(stones$any_special_anatomy == "Yes", na_rm = TRUE)
specialAnatFree <- n_perc0(stonesFree$any_special_anatomy == "Yes", na_rm = TRUE)
specialAnatResid <- n_perc0(stonesResid$any_special_anatomy == "Yes", na_rm = TRUE)
statSpecialAnat <- chisq.test(stones$any_special_anatomy, stones$residFrag, simulate.p.value = TRUE)
statspecialAnatPval <- round(statSpecialAnat$p.value, digits = 3)

### Left side ###
leftSide <- n_perc0(stones$side == "Left", na_rm = TRUE)
leftSideFree <- n_perc0(stonesFree$side == "Left", na_rm = TRUE)
leftSideResid <- n_perc0(stonesResid$side == "Left", na_rm = TRUE)
statLeftSide <- chisq.test(stones$side, stones$residFrag, simulate.p.value = TRUE)
statLeftSidePval <- round(statLeftSide$p.value, digits = 3)

### Right side ###
rightSide <- n_perc0(stones$side == "Right", na_rm = TRUE)
rightSideFree <- n_perc0(stonesFree$side == "Right", na_rm = TRUE)
rightSideResid <- n_perc0(stonesResid$side == "Right", na_rm = TRUE)
statRightSide <- chisq.test(stones$side, stones$residFrag, simulate.p.value = TRUE)
statRightSidePval <- round(statRightSide$p.value, digits = 3)

### Upper Pole
upper <- n_perc0(stones$upper == "Yes", na_rm = TRUE)
upperFree <- n_perc0(stonesFree$upper == "Yes", na_rm = TRUE)
upperResid <- n_perc0(stonesResid$upper == "Yes", na_rm = TRUE)
statUpper <- chisq.test(stones$upper, stones$residFrag, simulate.p.value = TRUE)
statUpperPval <- round(statUpper$p.value, digits = 3)

###  Lower Pole
lower <- n_perc0(stones$lowerPole == "Yes", na_rm = TRUE)
lowerFree <- n_perc0(stonesFree$lowerPole == "Yes", na_rm = TRUE)
lowerResid <- n_perc0(stonesResid$lowerPole == "Yes", na_rm = TRUE)
statLower <- chisq.test(stones$lowerPole, stones$residFrag, simulate.p.value = TRUE)
statLowerPval <- round(statLower$p.value, digits = 3)

###  Renal Pelvis
pelvis <- n_perc0(stones$renal_pelvis == "Yes", na_rm = TRUE)
pelvisFree <- n_perc0(stonesFree$renal_pelvis == "Yes", na_rm = TRUE)
pelvisResid <- n_perc0(stonesResid$renal_pelvis == "Yes", na_rm = TRUE)
statPelvis <- chisq.test(stones$renal_pelvis, stones$residFrag, simulate.p.value = TRUE)
statPelvisPval <- round(statPelvis$p.value, digits = 3)

###  Ureter
ureter <- n_perc0(stones$ureter == "Yes", na_rm = TRUE)
ureterFree <- n_perc0(stonesFree$ureter == "Yes", na_rm = TRUE)
ureterResid <- n_perc0(stonesResid$ureter == "Yes", na_rm = TRUE)
statUreter <- chisq.test(stones$ureter, stones$residFrag, simulate.p.value = TRUE)
statUreterPval <- round(statUreter$p.value, digits = 3)

### CaOx
comp_caox <- n_perc0(stones$caox == "Yes", na_rm = TRUE)
comp_caoxFree <- n_perc0(stonesFree$caox == "Yes", na_rm = TRUE)
comp_caoxResid <- n_perc0(stonesResid$caox == "Yes", na_rm = TRUE)
statComp_caox <- chisq.test(stones$caox, stones$residFrag, simulate.p.value = TRUE)
statComp_caoxPval <- round(statComp_caox$p.value, digits = 3)

### CaPhos
comp_caphos <- n_perc0(stones$caphos == "Yes", na_rm = TRUE)
comp_caphosFree <- n_perc0(stonesFree$caphos == "Yes", na_rm = TRUE)
comp_caphosResid <- n_perc0(stonesResid$caphos == "Yes", na_rm = TRUE)
statComp_caphos <- chisq.test(stones$caphos, stones$residFrag, simulate.p.value = TRUE)
statComp_caphosPval <- round(statComp_caphos$p.value, digits = 3)

### Uric Acid
comp_ua <- n_perc0(stones$uricacid == "Yes", na_rm = TRUE)
comp_uaFree <- n_perc0(stonesFree$uricacid == "Yes", na_rm = TRUE)
comp_uaResid <- n_perc0(stonesResid$uricacid == "Yes", na_rm = TRUE)
statComp_ua <- chisq.test(stones$uricacid, stones$residFrag, simulate.p.value = TRUE)
statComp_uaPval <- round(statComp_ua$p.value, digits = 3)

### Cystine
comp_cystine <- n_perc0(stones$cystine == "Yes", na_rm = TRUE)
comp_cystineFree <- n_perc0(stonesFree$cystine == "Yes", na_rm = TRUE)
comp_cystineResid <- n_perc0(stonesResid$cystine == "Yes", na_rm = TRUE)
statComp_cystine <- chisq.test(stones$cystine, stones$residFrag, simulate.p.value = TRUE)
statComp_cystinePval <- round(statComp_cystine$p.value, digits = 3)

### struvite
comp_struvite <- n_perc0(stones$struvite == "Yes", na_rm = TRUE)
comp_struviteFree <- n_perc0(stonesFree$struvite == "Yes", na_rm = TRUE)
comp_struviteResid <- n_perc0(stonesResid$struvite == "Yes", na_rm = TRUE)
statComp_struvite <- chisq.test(stones$struvite, stones$residFrag, simulate.p.value = TRUE)
statComp_struvitePval <- round(statComp_struvite$p.value, digits = 3)

### Make  the dataframe to feed the table ###
allData <- c(medianAge, men, meanWeight, meanBmi, medianASA, priorStone, specialAnat, "", leftSide, rightSide, "", upper, lower, pelvis, ureter, "", comp_caox, comp_caphos, comp_ua, comp_struvite, comp_cystine)
freeData <- c(medianAgeFree, menFree, meanWeightFree, meanBmiFree, medianASAFree, priorStoneFree, specialAnatFree, "", leftSideFree, rightSideFree, "", upperFree, lowerFree, pelvisFree, ureterFree, "", comp_caoxFree, comp_caphosFree, comp_uaFree, comp_struviteFree, comp_cystineFree)
residData <- c(medianAgeResid, menResid, meanWeightResid, meanBmiResid, medianASAResid, priorStoneResid, specialAnatResid, "", leftSideResid, rightSideResid, "", upperResid, lowerResid, pelvisResid, ureterResid, "", comp_caoxResid, comp_caphosResid, comp_uaResid, comp_struviteResid, comp_cystineResid)
pvals <- c(statAgePval, statMenPval, statWeightPval, statBmiPval, statASAPval, statPriorStonePval, statspecialAnatPval, "", statLeftSidePval, statRightSidePval, "", statUpperPval, statLowerPval, statPelvisPval, statUreterPval, "", statComp_caoxPval, statComp_caphosPval, statComp_uaPval, statComp_struvitePval, statComp_cystinePval)
t1 <- data.frame(row.names = c("Age, median (IQR), y", "Male, No. (%)", "Weight, mean (SD), kg", "BMI, mean (SD)", "ASA, median (IQR)", "Prior Stones, No.(%)", "Special GU Anatomy, No. (%)", "Side Of Treated Stone, No. (%)", "Left", "Right", "Stone Location (any portion of stone), No. (%)", "Upper Pole", "Lower Pole", "Renal Pelvis", "Ureter", "Stone Composition, No. (%)", "Calcium Oxalate", "Calcium Phosphate", "Uric Acid", "Struvite", "Cystine"), allData, freeData, residData, pvals)

### Create the table with Kable ###
kable(t1, booktabs = TRUE, caption = "Baseline Characteristics", col.names = c("All Patients (N = 50)", "Stone Free", "Residual Fragments", "P Value")) %>% 
  kable_styling() %>% 
  add_header_above(c("", "", "Results After PCNL" = 3)) %>% 
  group_rows("Patient Related", 1, 7) %>% 
  group_rows("Stone Related", 8, 21) %>% 
  add_indent(c(9, 10, 12, 13, 14, 15, 17, 18, 19, 20, 21))
```



```{r eval=FALSE, include=FALSE}
my_summary <- 
  list("Patient Related" =
         list("Age, median(IQR), y" = ~ qwraps2::median_iqr(age, na_rm = TRUE),
              "Male, No.(%)" = ~ qwraps2::n_perc0(gender == "M", na_rm = TRUE),
              "Weight, mean(SD), kg" = ~ qwraps2::mean_sd(weights, na_rm = TRUE)))
summary_table(stones, my_summary)
```


